### 2023-04-29 main/chapter_2ljets.tex
refer to new atlas fnp paper
incorporate Val's round 2 comments

% background diagrams
\begin{figure}[tp]
\centering
\begin{subfigure}{0.4\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_feynman_diboson_zz.png}
\caption{$ZZ$-like \diboson}
\end{subfigure}
\hfill
\begin{subfigure}{0.46\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_feynman_zyjets_slash.png}
\caption{$\zyjets$}
\end{subfigure}
\\
\begin{subfigure}{0.4\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_feynman_ttbar.png}
\caption{$\ttbar$}
\end{subfigure}
\hfill
\begin{subfigure}{0.46\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_feynman_top_decay.png}
\caption{Top quark decay}
\end{subfigure}
\caption[
Feynman diagrams for the Standard Model background processes
]{%
Feynman diagrams for the Standard Model background processes of important
background contributions in the $\twoljets$-electroweak search.
In \diboson\ events we expect one $Z$ boson-like decay to charged leptons, and
the other to neutrinos; additional jets can arise from initial state radiation
or additional strong radiations from the hard scatter.
Missing transverse momentum is absent from prompt emissions in $\zyjets$
events, as illustrated in (b), but arises from rare mismeasurements or
non-prompt decays.
Top quarks rapidly decay weakly, mainly to $b$-flavoured hadrons due to the
small weak mixing with the third generation;
panel (d) illustrates this with a leptonic decay that we expect to select.
}
\label{fig:2ljets_feynman}
\end{figure}

@@

Signal regions of the $\twoljets$-electroweak analysis are combined to form
five discovery regions.
Results of their model-independent interpretations are presented in
Table~\ref{tab:2ljets_discovery}.
The same data deficits drive stronger bounds on possible signals that manifest
as smaller limits on the signal expectations, which are labelled
$S_{\mathrm{obs}}^{95}$ in the figure.
\footnotetext{%
The visible cross-section is defined such that the expected event count is
given by the product of $A\epsilon\sigma\times L$, where $L$ is the integrated
luminosity (typically in $\mathrm{fb}^{-1}$ units).
This particular presentation is uses an `acceptance' $A$, an
`efficiency' $\epsilon$, and an observable cross-section $\sigma_{\mathrm{obs}}$
(presented here at at $95\%$ limit).
The acceptance and efficiency are defined using truth-level simulations, in
which the $A$ is the fraction of truth-level events that pass selection
requirements, and $\epsilon$ is the ratio from the truth-level yield to the
reconstruction-level yield.
Because detector effects can increase yields, this $\epsilon$ can exceed
one~\cite{susy_results_presentation_papers}.
}

@@

Leptons in $\twoljets$ are further sliced into two classes:
`baseline' and `signal'.
Baseline leptons and are used in background estimation methods and to ensure
orthogonality with other \atlas\ analyses that study events with differing
lepton counts, such that our results can be combined in future to deliver an
improved total sensitivity.
Baseline requirements are softer and lower quality than signal requirements;
the signal leptons are a subset of the baseline leptons.
Therefore if a lepton is a signal lepton, then it is also a baseline lepton,
but for clarity we often prefer to state that both requirements are met.

@@

The power of each of these variables to select signals and exclude backgrounds
is hugely variable across our many signal models and regions of phase space.
For illustration, Table~\ref{tab:2ljets_ewk_acceff} reproduces some signal
efficiency numbers for the sum of yields in all $\twoljets$-electroweak signal
regions.
It finds a that only $0.0003\%$ of signal events are captured under the light
C1N2 model with
$m(\chargino_1, \neutralino_2) = 100\,\eV[G]$ and
$m(\neutralino_1) = 60\,\eV[G]$.
The reduced cross-sections for models with large masses are offset by their
more spectacular appearances, with a much larger capture rate of
for more extreme models.
For example, the corresponding capture rate it $0.64\%$ for the C1N2 model with
$m(\chargino_1, \neutralino_2) = 650\,\eV[G]$ and
$m(\neutralino_1) = 250\,\eV[G]$.
Detailed cut-flow tables,
Tables~\ref{aux:ewk:cutflowSRllbb} to~\ref{aux:ewk:cutflowSROffShell_2} of
Appendix~\ref{app:2ljets_cut_flows},
give more granular information into the effects of many intermediate
selection requirements across all signal regions for several representative
signal models.
Notably, $\mttwoll$ rejects substantial amounts of signal, although this
rejection is worthwhile for its larger rate of background rejection.

On background rejection performance,
Tables~\ref{tab:cutflows_CR-VZ} to~\ref{tab:cutflows_CR-DY} of
Appendix~\ref{app:2ljets_cut_flows} present cut-flow tables in the four control
regions of $\twoljets$-electroweak with background sample rates included.
These tables show that $\mttwoll$ selections are powerful for removing $\ttbar$
backgrounds, as expected, and that $\metsig$ selections beyond the
pre-selection requirement of $\metsig > 6$ drastically reduce most other
backgrounds.
Summed across all four control regions with data given in these tables,
the control region selections reduce background rates with a survival fraction
of $1.5\%$ relative to the pre-selection requirements of
Table~\ref{tab:2ljets_presel}.

@@

\subsubsection{Low-ISR}
\label{sec:2ljets_low_isr}
The Low regions presented in Section~\ref{sec:2ljets_low} resulted from a
hasty redesign shortly before unblinding.
Before that redesign, a multi-jet signal region SR-Low-ISR was included,
with associated validation and control regions
that required $\njet \geq 3$ and offered some meagre extra sensitivity.
Here, `ISR' stands for Initial-State Radiation, following the naming
of previous \atlas\ searches including the partial Run~2 $2\ell$/$3\ell$
search, which we are updating~\cite{atlas_23l_SUSY_2016_24} in the
$\twoljets$-electroweak search.
Although these Low-ISR regions were removed from the analysis and we never
inspected the data in SR-Low-ISR (never unblinded), I include them here because
they reveal some interesting context to the surviving analysis design.
Specifically, prototype validation regions for Low-ISR designs revealed some
bad modelling by our Standard Model background samples in phase space that is
close to SR-Low, which is a region where we observe significant data deficits.
Secondly, these regions were an impactful part of the analysis design process
and led to the final design of SR-Low;
describing my experiences with them therefore reveals more about the true
design process.

@@

Yes, actual data are histogrammed by the Matrix Method to assign its fake
estimates, and those data include tight-tight events that enter our search
regions --- the very same data in which we are searching for anomalous signals.
Since it includes actual data, this method unblinds all regions,
including signal regions, early.
And because it knows and uses the data we are supposed to be predicting, it is
no prediction at all!

@@

This formulation of extrapolation as a weighted mixture has some freedom in
which to assign the distance function $d(\vec x, \vec y)$, and within that
freedom we can make choices that encode some physical understanding.
In the case of our C1N2 model, two characteristic features are the
parent mass $m(\chargino_1, \neutralino_2) = m_a$
and the mass splitting $m_a - m_b$, and we want an extrapolation scheme which
aligns changes in these variables with changes in the extrapolated value.
From some experimentation, I found that a Manhattan distance in these variables
implements such attractive extrapolation:
being a linear sum, the exponentiated Manhattan distance beneficially
separates into a product of terms, so contours of constant $r_j(\vec x)$ are
segmented into straight lines that relate to the chosen coordinates.
Alternatives, such as the Euclidean distance, produced curved contours that did
not align as well with the $m_a$ and $m_a - m_b$ directions.
Since the mass distance is dimensionful, it must be scaled to a reference
value, which I chose as $100\,\eV[G]$; therefore
\begin{equation}
d_\mathrm{C1N2}(\vec x, \vec y) =
\frac{1}{100\,\eV[G]}
\big(
|m_a^{\vec x} - m_b^{\vec y}|
+
|(m_a^{\vec x} - m_b^{\vec x}) - (m_a^{\vec y} - m_b^{\vec y})|
\big)
.
\end{equation}
To put the Higgs branching fraction $B$ on a similar scale to the parent mass
for GMSB models, its distance measure has a separate scale factor;
\begin{equation}
d_\mathrm{GMSB}(\vec x, \vec y) =
\frac{1}{100\,\eV[G]}
|m_a^{\vec x} - m_b^{\vec y}|
+
5
|B^{\vec x} - B^{\vec y}|
.
\end{equation}
Some example extrapolations according to these rules are displayed in
Figure~\ref{fig:2ljets_signal_sys_extrap}.
This figure illustrates the common theme that our theory variations are small
if and only if signal yields are large enough for good statical precision.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figures/2ljets_summary_no_sr_log.png}
\caption[
Alternative summary table with backgrounds fitted to control regions only
]{%
Alternative summary table with backgrounds fitted to control regions only.
A version fitted also to signal regions is shown in
Figure~\ref{fig:2ljets_summary}.
Unlike Figure~\ref{fig:2ljets_summary} the this figure does not adjust its
background estimates to match the observed data in signal regions; notably,
this figure has much larger background predictions in SR-Low-a and SR-Low-b,
with correspondingly larger negative significances in those regions.
The two figures also have aesthetic differences in the colouring and grouping
of background samples.
\\[0.4em]
Fitting to signal regions is the policy of the \atlas\ SUSY group, but deviates
from the intentions of the significance measure.
Because the post-fit yields in SR-Low achieve a small negative significance,
all mentions of its data deficit were removed from the paper.
}
\label{fig:2ljets_summary_no_sr}
\end{figure}

@@

Collision energies ultimately bound the masses of heavy particles that can
be produced in collisions, so stronger accelerators will be necessary
to expand the search in that direction.
Future prospects for the coming decades at CERN include the
High-Energy LHC (HE-LHC), which could achieve $\sqrt{s}=27\,\eV[T]$ with new
hardware in the current LHC tunnel~\cite{FCC:2018byv},
and a Future Circular Collider (FCC), whose new $100\,\mathrm{km}$
`quasi-circular' tunnel would permit proton collisions at
$\sqrt{s}=100\,\eV[T]$~\cite{FCC:2018vvp}.
Just as the LHC follows the LEP electron-positron collider, this FCC hadron
collider (FCC-hh) might be proceeded by an electron-positron collider (FCC-ee)
with centre-of-mass energies high enough (up to $\sqrt{s}=350\,\eV[T]$) to
pair-produce all Standard Model particles~\cite{FCC:2018evy}.
Such scrutiny of the Standard Model is also a search for new physical effects,
and including some that could be explained with supersymmetric
models~\cite{FCC:2018byv, Gounaris1999}.

>>>>

% background diagrams
\begin{figure}[tp]
\centering
\begin{subfigure}{0.4\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_feynman_diboson_zz.png}
\caption{$ZZ$-like \diboson}
\end{subfigure}
\hfill
\begin{subfigure}{0.46\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_feynman_zyjets_slash.png}
\caption{$\zyjets$}
\end{subfigure}
\\
\begin{subfigure}{0.4\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_feynman_ttbar.png}
\caption{$\ttbar$}
\end{subfigure}
\hfill
\begin{subfigure}{0.46\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/2ljets_feynman_top_decay.png}
\caption{Top quark decay}
\end{subfigure}
\caption[
Example Feynman diagrams for the Standard Model background processes
]{%
Example Feynman diagrams for the Standard Model background processes of
important background contributions in the $\twoljets$-electroweak search.
In \diboson\ events we expect one $Z$ boson-like decay to charged leptons, and
the other to neutrinos; additional jets can arise from initial state radiation
or additional strong radiations from the hard scatter.
Missing transverse momentum is absent from prompt emissions in $\zyjets$
events, as illustrated in (b), but arises from rare mismeasurements or
non-prompt decays.
Top quarks rapidly decay weakly, mainly to $b$-flavoured hadrons due to the
small weak mixing with the third generation;
panel (d) illustrates this with a leptonic decay that we expect to select.
}
\label{fig:2ljets_feynman}
\end{figure}

@@

Signal regions of the $\twoljets$-electroweak analysis are combined to form
five discovery regions.
Results of their model-independent interpretations are presented in
Table~\ref{tab:2ljets_discovery}.
The same data deficits drive stronger bounds on possible signals that manifest
as smaller limits on the signal expectations, which are labelled
$S_{\mathrm{obs}}^{95}$ in this table.
\footnotetext{%
The visible cross-section is defined such that the expected event count is
given by the product of $A\epsilon\sigma\times L$, where $L$ is the integrated
luminosity (typically in $\mathrm{fb}^{-1}$ units).
This particular presentation is uses an `acceptance' $A$, an
`efficiency' $\epsilon$, and an observable cross-section $\sigma_{\mathrm{obs}}$
(presented here at at $95\%$ limit).
The acceptance and efficiency are defined using truth-level simulations, in
which the $A$ is the fraction of truth-level events that pass selection
requirements, and $\epsilon$ is the ratio from the truth-level yield to the
reconstruction-level yield.
Because detector effects can increase yields, this $\epsilon$ can exceed
one~\cite{susy_results_presentation_papers}.
}

@@

Leptons in $\twoljets$ are further sliced into two classes:
`baseline' and `signal'.
Baseline leptons and are used in background estimation methods and to ensure
orthogonality with other \atlas\ analyses that study events with differing
lepton counts, such that our results can be combined in future to deliver an
improved total sensitivity.
Baseline requirements are softer and lower quality than signal requirements;
the signal leptons are a subset of the baseline leptons.
Therefore, if a lepton is a signal lepton, then it is also a baseline lepton,
but for clarity we often prefer to state that both requirements are met.

@@

The power of each of these variables to select signals and exclude backgrounds
is hugely variable across our many signal models and regions of phase space.
For illustration, Table~\ref{tab:2ljets_ewk_acceff} reproduces some signal
efficiency numbers for the sum of yields in all $\twoljets$-electroweak signal
regions.
It finds a that only $0.0003\%$ of signal events are captured under the light
C1N2 model with
$m(\chargino_1, \neutralino_2) = 100\,\eV[G]$ and
$m(\neutralino_1) = 60\,\eV[G]$.
The reduced cross-sections for models with large masses are offset by their
more spectacular appearances, with a much larger capture rate of
for more extreme models.
For example, the corresponding capture rate is $0.64\%$ for the C1N2 model with
$m(\chargino_1, \neutralino_2) = 650\,\eV[G]$ and
$m(\neutralino_1) = 250\,\eV[G]$.
Detailed cut-flow tables,
Tables~\ref{aux:ewk:cutflowSRllbb} to~\ref{aux:ewk:cutflowSROffShell_2} of
Appendix~\ref{app:2ljets_cut_flows},
give more granular information into the effects of many intermediate
selection requirements across all signal regions for several representative
signal models.
Notably, $\mttwoll$ rejects substantial amounts of signal, although this
rejection is worthwhile for its larger rate of background rejection.

On background rejection performance,
Tables~\ref{tab:cutflows_CR-VZ} to~\ref{tab:cutflows_CR-DY} of
Appendix~\ref{app:2ljets_cut_flows} present cut-flow tables in the four control
regions of $\twoljets$-electroweak, with background sample rates included.
These tables show that $\mttwoll$ selections are powerful for removing $\ttbar$
backgrounds, as expected, and that $\metsig$ selections beyond the
pre-selection requirement of $\metsig > 6$ drastically reduce most other
backgrounds.
Summed across all four control regions with data given in these tables,
the control region selections reduce background rates with a survival fraction
of $1.5\%$ relative to the pre-selection requirements of
Table~\ref{tab:2ljets_presel}.

@@

\subsubsection{Low-ISR}
\label{sec:2ljets_low_isr}
The Low regions presented in Section~\ref{sec:2ljets_low} resulted from a
hasty redesign shortly before unblinding.
Before that redesign, a multi-jet signal region SR-Low-ISR was included,
with associated validation and control regions
that required $\njet \geq 3$ and offered some meagre extra sensitivity.
Here, `ISR' stands for Initial-State Radiation, following the naming
of previous \atlas\ searches including the partial Run~2 $2\ell$/$3\ell$
search, which we are updating~\cite{atlas_23l_SUSY_2016_24} in the
$\twoljets$-electroweak search.
Although these Low-ISR regions were removed from the analysis and we never
inspected the data in SR-Low-ISR (this signal region was never unblinded),
I include them here because they reveal some interesting context to the
surviving analysis design.
Specifically, prototype validation regions for Low-ISR designs revealed some
bad modelling by our Standard Model background samples in phase space that is
close to SR-Low, which is a region where we observe significant data deficits.
Secondly, these regions were an impactful part of the analysis design process
and led to the final design of SR-Low;
describing my experiences with them therefore reveals more about the true
design process.

@@

Yes, actual data are histogrammed by the Matrix Method to assign its fake
estimates, and those data include tight-tight events that enter our search
regions --- the very same data in which we are searching for anomalous signals.
Since it includes actual data, this method unblinds all regions,
including signal regions, early~\cite{ATLAS:2022swp}.
And because it knows and uses the data we are supposed to be predicting, it is
no prediction at all!

@@

This formulation of extrapolation as a weighted mixture has some freedom in
which to assign the distance function $d(\vec x, \vec y)$, and within that
freedom we can make choices that encode some physical understanding.
In the case of our C1N2 model, two characteristic features are the
parent mass $m(\chargino_1, \neutralino_2) = m_a$
and the mass splitting $m_a - m_b$, and we want an extrapolation scheme that
aligns changes in these variables with changes in the extrapolated value.
From some experimentation, I found that a Manhattan distance in these variables
implements such attractive extrapolation:
being a linear sum, the exponentiated Manhattan distance beneficially
separates into a product of terms, so contours of constant $r_j(\vec x)$ are
segmented into straight lines that relate to the chosen coordinates.
Alternatives, such as the Euclidean distance, produced curved contours that did
not align as well with the $m_a$ and $m_a - m_b$ directions.
Since the mass distance is dimensionful, it must be scaled to a reference
value, which I chose as $100\,\eV[G]$; therefore
\begin{equation}
d_\mathrm{C1N2}(\vec x, \vec y) =
\frac{1}{100\,\eV[G]}
\big(
|m_a^{\vec x} - m_b^{\vec y}|
+
|(m_a^{\vec x} - m_b^{\vec x}) - (m_a^{\vec y} - m_b^{\vec y})|
\big)
.
\end{equation}
To put the Higgs branching fraction $B$ on a similar scale to the parent mass
for GMSB models, its distance measure has a separate scale factor;
\begin{equation}
d_\mathrm{GMSB}(\vec x, \vec y) =
\frac{1}{100\,\eV[G]}
|m_a^{\vec x} - m_b^{\vec y}|
+
5
|B^{\vec x} - B^{\vec y}|
.
\end{equation}
Some example extrapolations according to these rules are displayed in
Figure~\ref{fig:2ljets_signal_sys_extrap}.
This figure illustrates the common theme that our theory variations are small
if and only if signal yields are large enough for good statical precision.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figures/2ljets_summary_no_sr_log.png}
\caption[
Alternative summary table with backgrounds fitted to control regions only
]{%
Alternative summary table with backgrounds fitted to control regions only.
A version fitted also to signal regions is shown in
Figure~\ref{fig:2ljets_summary}.
Unlike Figure~\ref{fig:2ljets_summary}, this figure does not adjust its
background estimates to match the observed data in signal regions; notably,
this figure has much larger background predictions in SR-Low-a and SR-Low-b,
with correspondingly larger negative significances in those regions.
The two figures also have aesthetic differences in the colouring and grouping
of background samples.
\\[0.4em]
Fitting to signal regions is the policy of the \atlas\ SUSY group, but deviates
from the intentions of the significance measure.
Because the post-fit yields in SR-Low achieve a small negative significance,
all mentions of its data deficit were removed from the paper.
}
\label{fig:2ljets_summary_no_sr}
\end{figure}

@@

Collision energies ultimately bound the masses of heavy particles that can
be produced in collisions, so stronger accelerators will be necessary
to expand the search in that direction.
Future prospects for the coming decades at CERN include the
High-Energy LHC (HE-LHC), which could achieve $\sqrt{s}=27\,\eV[T]$ with new
hardware in the current LHC tunnel~\cite{FCC:2018byv},
and a Future Circular Collider (FCC), whose new $100\,\mathrm{km}$
`quasi-circular' tunnel would permit proton collisions at
$\sqrt{s}=100\,\eV[T]$~\cite{FCC:2018vvp}.
Just as the LHC follows the LEP electron-positron collider, this FCC hadron
collider (FCC-hh) might be preceeded by an electron-positron collider (FCC-ee)
with centre-of-mass energies high enough (up to $\sqrt{s}=350\,\eV[G]$) to
pair-produce all Standard Model particles~\cite{FCC:2018evy}.
Such scrutiny of the Standard Model is also a search for new physical effects,
and including some that could be explained with supersymmetric
models~\cite{FCC:2018byv, Gounaris1999}.
